### 2020-03-25 CalNotes.tex
Partial update in restructuring the paper

In the rest of this section we will focus on the properties of the process $Z^\epl$ when it is considered together with the original process $X^\epl$. In particular, we focus on ergodic properties and multiscale convergence. Finally, we present and analyse unbiased estimators obtained with the filtering process.

\subsection{Ergodic properties of the filter}\label{sec:ergodic}

Let us consider the filtering kernel \eqref{eq:filter} with $\beta = 1$, i.e.,
\begin{equation}\label{eq:filter_beta1}
	k(r) = \frac{1}{\delta} e^{-r/\delta}.
\end{equation}
In this case, Leibniz integral rule yields the equality
\begin{equation}
	\d Z^\epl_t = k(0) X^\epl_t \dd t + \int_0^t k'(t-s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t,
\end{equation}
which can be interpreted as an ordinary differential equation for $Z_t^\epl$ driven by the stochastic signal $X^\epl$. Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain the system of two one-dimensional SDEs
\begin{equation}
\label{eq:systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) \dd t+ \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
The first ingredient for verifying the ergodic properties of the two-dimensional process $(X^\epl, Z^\epl)^\top \defeq ((X^\epl_t, Z^\epl_t)^\top, 0 \leq t \leq T)$ is verifying that the measure induced by the stochastic process admits a smooth density with respect to the Lebesgue measure. Since noise is present only on the first component, this is a consequence of the theory of hypo-ellipticity, as summarized in the following Lemma, \corr{whose proof is given in Appendix \ref{ap:Proofs}.}

@@

Once it is established that the law of the process admits a smooth density for all times $t>0$, which satisfies a time-dependent Fokker--Planck equation, we are interested in the limiting properties of this law. In particular, we know that the process $X^\epl$ alone is geometrically ergodic \cite[Theorem 4.4]{MSH02}, and we wish the couple $(X^\epl, Z^\epl)^\top$ to be endowed with the same property. The following Lemma guarantees that the couple is indeed geometrically ergodic, \corr{and its proof is given in Appendix \ref{ap:Proofs}.}

@@

In a general case, it is not possible to find an explicit solution to \eqref{eq:FPsystem}. Nevertheless, it is possible to show some relevant properties of the solution itself, which are summarized in the following Lemma, \corr{whose proof is given in Appendix \ref{ap:Proofs}.}

@@

\subsection{Multiscale convergence}\label{sec:convMS}

We now investigate the convergence of the couple $(X^\epl, Z^\epl)^\top$ with respect to the multiscale parameter $\epl \to 0$. In particular, it is known that the invariant measure of $X^\epl$ converges weakly to the invariant measure of $X$ solution of the homogenized equation \eqref{eq:SDE_HOM}. The following result, \corr{whose proof is given in Appendix \ref{ap:Proofs},} guarantees the same kind of convergence for the couple $(X^\epl, Z^\epl)^\top$.

\begin{lemma}\label{lem:convMeasure} Under Assumption \ref{as:regularity}, let $\mu^\epl$ be the invariant measure of the couple $(X^\epl, Z^\epl)^\top$ and let $(X_0^\epl, Z_0^\epl)^\top \sim \mu^\epl$. Then, the measure $\mu^\epl$ converges weakly to the measure $\mu^0(\d x, \d z) = \rho^0(x, z) \dd x \dd z$, whose density $\rho^0$ is the unique solution of the Fokker--Planck equation
\begin{equation} \label{eq:FPsystem_homogenized}
	\Sigma \partial^2_{xx} \rho^0(x,z) + \partial_x\left( A \cdot V'(x) \rho^0(x,z) \right) + \frac{1}{\delta}\partial_z\left((z - x) \rho^0(x,z) \right) = 0,
\end{equation}
where $A$ and $\Sigma$ are the coefficients of the homogenized equation \eqref{eq:SDE_HOM}.
\end{lemma}

@@

\subsection{The filtering-based estimator}\label{sec:FilterMLE}

We now consider the inference problem and propose our filtering-based estimator of the drift, which is formally given by the formula
\begin{equation}\label{eq:AHatMixed}
	\widehat A^\epl_k(T) = - \widetilde M^{-1} \widetilde h,
\end{equation}
where we employ the subscript $k$ for reference to the filter's kernel in \eqref{eq:filter}, and where
\begin{equation}\label{eq:MandHTilde}
	\widetilde M = \frac{1}{T} \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t, \qquad \text{and} \qquad \widetilde h = \frac{1}{T} \int_0^T V'(Z^\epl_t) \dd X^\epl_t.
\end{equation}
Let us remark that the formula above is obtained from \eqref{eq:MLE} by replacing only one instance of $X_t^\epl$ with $Z_t^\epl$ in both $M$ and $h$. In particular, it is fundamental for proving unbiasedness to keep in the definition of $h$ the differential of the original process $\d X^\epl_t$. Let us furthermore remark that $\widehat A^\epl_k(T)$ need not be the minimizer of some filtering-based likelihood function. In fact, if one were to replace $Z_t^\epl$ directly in \eqref{eq:Likelihood}, the symmetric part of the matrix $\widetilde M$ would appear and $\widehat A^\epl_k(T)$ would not be the minimizer. Therefore, the estimator $\widehat A^\epl_k(T)$ has to be thought of as a perturbation of $\widehat A^\epl(T)$, directly at the level of estimators and after the maximization procedure. The only theoretical guarantee which is still needed for the well-posedness of $\widehat A^\epl_k(T)$ is for $\widetilde M$ to be invertible, which we assume to be true and which we observed to hold in practice.

%\begin{lemma}\label{lem:AHatWellPosed} Under Assumption \ref{as:regularity}, the matrix $\widetilde M$ defined in \eqref{eq:MandHTilde} is invertible a.s. 
%\end{lemma}
%\begin{proof} 
%\todo: \corr{Idea. We don't need $\widetilde M$ to be symmetric positive definite (it is not) due to the comment just above the lemma. Nevertheless, if the functions $V_i'(x)$ are not such that the rows of $\widetilde M$ are identical for any $(X_t, Z_t)$, then the randomness in the entries should guarantee that $\det \widetilde M \neq 0$ with probability one. Still missing the (probably trivial) argument why this is true.}
%\end{proof}

We are now able to present the main result of unbiasedness of the filtering-based estimator.
\begin{theorem}\label{thm:mainTheorem} Let the assumptions of Lemma \ref{lem:ergodicity} and Lemma \ref{lem:convMeasure} hold, and let $\widehat A^\epl_k(T)$ be defined in \eqref{eq:AHatMixed}. If $\widetilde M$ is invertible, then
	\begin{equation}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A_k^\epl(T) = A, \quad \text{a.s.},
	\end{equation}
	where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
\end{theorem}

@@

\lim_{T \to \infty} I_1^\epl(T) = \E^{\rho^\epl} [V'(Z^\epl) \otimes V'(X^\epl)]^{-1} \E^{\rho^\epl} \left [ \frac{1}{\epl} p' \left ( \frac{X^\epl}{\epl} \right ) V'(Z^\epl) \right ], \quad \text{a.s.}

@@

	&= -  \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ]\alpha + \sigma \int_{\R} \int_{\R} V'(z) \phi^\epl(x) \psi^\epl(z) \partial_x R^\epl(x,z) \dd x \dd z.

@@

\lim_{T \to \infty} I_1^\epl(T) = -\alpha + \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ]^{-1} \sigma \int_{\R} \int_{\R} V'(z) \phi^\epl(x) \psi^\epl(z) \partial_x R^\epl(x,z) \dd x \dd z, \quad \text{a.s.}

@@

	\lim_{T \to \infty} I_1^\epl(T) = -\alpha + \frac1\delta \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ]^{-1} \E^{\rho^\epl}[(X^\epl - Z^\epl)^2 V''(Z^\epl)], \quad \text{a.s.}	

@@

We now pass to the limit as $\epl$ goes to zero and
\begin{equation} \label{limit_estimator_2}
	\lim_{\epl \to 0} \lim_{T \to \infty} I_1^\epl(T) = -\alpha + \frac1\delta \E^{\rho^0} [ V'(Z) \otimes V'(X) ]^{-1} \E^{\rho^0}[(X - Z)^2 V''(Z)], \quad \text{a.s.},

@@

where the function $\rho^0$ is the density of the limit invariant distribution for $\epl \to 0$, as in Lemma \ref{lem:convMeasure}. Due to Corollary \ref{lem:FPMarginal_Hom}, we have

@@

	&=  \E^{\rho^0} [ V'(Z) \otimes V'(X) ]A.

@@

	\lim_{T \to \infty} I_{2,1}^\epl(T) = \E^{\rho^\epl}\left[V'(Z^\epl) \otimes V'(X^\epl)\right]^{-1}\E^{\rho^\epl}\left[V'(Z^\epl) \otimes V'(Z^\epl)\right] \eqdef R^\epl,

@@

\corr{ CASE $\delta = \epl^\zeta$ (PROOF IN APPENDIX \ref{ap:zeta}).
\begin{theorem}\label{thm:mainTheorem_zeta} Let the assumptions of Lemma \ref{lem:ergodicity} and Lemma \ref{lem:convMeasure} hold, let $\widehat A^\epl_k(T)$ be defined in \eqref{eq:AHatMixed} and let $\delta = \epl^\zeta$ with $\zeta \in (0,2)$. If $\widetilde M$ is invertible, then
		\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A_k^\epl(T) = A, \quad \text{in law},
		\end{equation}
		where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
\end{theorem}}

@@

\section{Proof of Lemmas}\label{ap:Proofs}

@@

	which concludes the proof.

Corollary \ref{cor:distanceZandX} is a direct consequence of Lemma \ref{lem:zeta} and provides a rough estimate of the difference between the trajectories $X_t^\epl$ and $Z_t^\epl$ when they are at stationarity.

>>>>

\begin{figure}[t]
	\centering
	\begin{tabular}{c}
		\includegraphics[]{Figures/TrajD1} \\
		\includegraphics[]{Figures/TrajD5} \\
		\includegraphics[]{Figures/TrajD10}
	\end{tabular}
	\caption{\corr{Filtering a trajectory $X^\epl$ obtained with $V(x) = x^2 / 2$, $p(y) = \cos(y)$, $\alpha = 1$, $\sigma = 0.5$ and $\epl = 0.1$. The filtering width is $\delta = \{1, \sqrt{\epl}, \epl\}$ from top to bottom, respectively.}}
	\label{fig:filter}
\end{figure}

\corr{Given a trajectory $X^\epl$ and the filtered data $Z^\epl$, the estimator of the drift coefficient we propose is given by
\begin{equation}\label{eq:AHatMixed}
	\widehat A^\epl_k(T) = - \widetilde M^{-1} \widetilde h,
\end{equation}
where we employ the subscript $k$ for reference to the filter's kernel in \eqref{eq:filter}, and where
\begin{equation}\label{eq:MandHTilde}
\widetilde M = \frac{1}{T} \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t, \qquad \text{and} \qquad \widetilde h = \frac{1}{T} \int_0^T V'(Z^\epl_t) \dd X^\epl_t.
\end{equation}
Let us remark that the formula above is obtained from \eqref{eq:MLE} by replacing only one instance of $X_t^\epl$ with $Z_t^\epl$ in both $M$ and $h$. In particular, it is fundamental for proving unbiasedness to keep in the definition of $h$ the differential of the original process $\d X^\epl_t$. Let us furthermore remark that $\widehat A^\epl_k(T)$ need not be the minimizer of some filtering-based likelihood function. In fact, if one were to replace $Z_t^\epl$ directly in \eqref{eq:Likelihood}, the symmetric part of the matrix $\widetilde M$ would appear and $\widehat A^\epl_k(T)$ would not be the minimizer. Therefore, the estimator $\widehat A^\epl_k(T)$ has to be thought of as a perturbation of $\widehat A^\epl(T)$, directly at the level of estimators and after the maximization procedure. The only theoretical guarantee which is still needed for the well-posedness of $\widehat A^\epl_k(T)$ is for $\widetilde M$ to be invertible, which we assume to be true and which we observed to hold in practice.

Let us from now on consider $\beta = 1$. For this value of $\beta$, the parameter $\delta$ appearing in \eqref{eq:filter} regulates the width of the filtering window. In practice, larger values of $\delta$ will lead to trajectories which are smoother and for which fast-scale oscillations are practically canceled. Let us remark that the filtering width resembles the subsampling step employed for the estimator $\widehat A^\epl_\delta$ introduced and analyzed in \cite{PaS07}. For subsampling, the choice guaranteeing asymptotically unbiased results is $\delta = \epl^\zeta$ with $\zeta \in (0, 1)$, and a similar analysis is due for our filtering technique. For visualization purposes, we depict in Figure \ref{fig:filter} the filtered trajectory $Z^\epl$ for three different values of $\delta$, namely $\delta = \{1, \sqrt{\epl}, \epl\}$. With $\delta = 1$, all oscillations at the fast scale are canceled and the filtered trajectory $Z^\epl$ presents only slow-scale variations are present in $Z^\epl$. Reducing the value of $\delta$, fast-scale oscillations are progressively taken into account.

In the following, we first focus on the ergodic properties of the process $Z^\epl$ when it is coupled with the process $X^\epl$. This analysis is practically independent of the choice of $\delta$, and is therefore presented on its own. Then, we focus on two different cases which depend on the choice of the width $\delta$ of the filter. First, we consider $\delta$ to be of the magnitude of the slow-scale variations, i.e., $\delta \propto \epl^0$. In this case, an analysis based on multiscale convergence is presented. We then move on to the case $\delta \propto \epl^\zeta$, and show how to choose the exponent $\zeta$ in order to obtain unbiased estimators. For this second case, we widely employ techniques and estimates which come from \cite{PaS07}.
}

\subsection{Ergodic properties}\label{sec:ergodic}

Let us consider the filtering kernel \eqref{eq:filter} with $\beta = 1$, i.e.,
\begin{equation}\label{eq:filter_beta1}
	k(r) = \frac{1}{\delta} e^{-r/\delta}.
\end{equation}
In this case, Leibniz integral rule yields the equality
\begin{equation}
	\d Z^\epl_t = k(0) X^\epl_t \dd t + \int_0^t k'(t-s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t,
\end{equation}
which can be interpreted as an ordinary differential equation for $Z_t^\epl$ driven by the stochastic signal $X^\epl$. Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain the system of two one-dimensional SDEs
\begin{equation}
\label{eq:systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) \dd t+ \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
The first ingredient for verifying the ergodic properties of the two-dimensional process $(X^\epl, Z^\epl)^\top \defeq ((X^\epl_t, Z^\epl_t)^\top, 0 \leq t \leq T)$ is verifying that the measure induced by the stochastic process admits a smooth density with respect to the Lebesgue measure. Since noise is present only on the first component, this is a consequence of the theory of hypo-ellipticity, as summarized in the following Lemma, whose proof is given in Appendix \ref{ap:Proofs}.

@@

Once it is established that the law of the process admits a smooth density for all times $t>0$, which satisfies a time-dependent Fokker--Planck equation, we are interested in the limiting properties of this law. In particular, we know that the process $X^\epl$ alone is geometrically ergodic \cite[Theorem 4.4]{MSH02}, and we wish the couple $(X^\epl, Z^\epl)^\top$ to be endowed with the same property. The following Lemma guarantees that the couple is indeed geometrically ergodic, and its proof is given in Appendix \ref{ap:Proofs}.

@@

In a general case, it is not possible to find an explicit solution to \eqref{eq:FPsystem}. Nevertheless, it is possible to show some relevant properties of the solution itself, which are summarized in the following Lemma, whose proof is given in Appendix \ref{ap:Proofs}.

@@

\subsection{Filtering at the slow scale}\label{sec:Slow}

\corr{In this section, we analyze the behavior of the filtering-based estimator $\widehat A^\epl_k(T)$ given in \eqref{eq:AHatMixed} when the filtering width $\delta$ is independent of $\epl$. In particular, we suppose there exist positive constants $C_1$ and $C_2$ independent of $\epl$ and $\delta$ such that $C_1 \epl^0 \leq \delta \leq C_2 \epl^0$. In the following, we will write more compactly that $\delta = \mathcal O(\epl^0)$. The analysis in this case is based on the convergence of the couple $(X^\epl, Z^\epl)^\top$ with respect to the multiscale parameter $\epl \to 0$.} In particular, it is known that the invariant measure of $X^\epl$ converges weakly to the invariant measure of $X$ solution of the homogenized equation \eqref{eq:SDE_HOM}. The following result, whose proof is given in Appendix \ref{ap:Proofs}, guarantees the same kind of convergence for the couple $(X^\epl, Z^\epl)^\top$.

\begin{lemma}\label{lem:convMeasure} Under Assumption \ref{as:regularity}, let $\mu^\epl$ be the invariant measure of the couple $(X^\epl, Z^\epl)^\top$ and let $(X_0^\epl, Z_0^\epl)^\top \sim \mu^\epl$. \corr{If $\delta = \mathcal O(\epl^0)$,} then the measure $\mu^\epl$ converges weakly to the measure $\mu^0(\d x, \d z) = \rho^0(x, z) \dd x \dd z$, whose density $\rho^0$ is the unique solution of the Fokker--Planck equation
\begin{equation} \label{eq:FPsystem_homogenized}
	\Sigma \partial^2_{xx} \rho^0(x,z) + \partial_x\left( A \cdot V'(x) \rho^0(x,z) \right) + \frac{1}{\delta}\partial_z\left((z - x) \rho^0(x,z) \right) = 0,
\end{equation}
where $A$ and $\Sigma$ are the coefficients of the homogenized equation \eqref{eq:SDE_HOM}.
\end{lemma}

\begin{proof} Let $(X, Z)^\top \defeq \left((X_t, Z_t)^\top, 0\leq t \leq T\right)$ be the solution of
	\begin{equation}
	\label{eq:systemSDEHom}
	\begin{aligned}
	\d X_t &= -A \cdot V'(X_t) \dd t + \sqrt{2\Sigma} \dd W_t, \\
	\d Z_t &= \frac{1}{\delta} \left ( X_t - Z_t \right ) \dd t,
	\end{aligned}
	\end{equation} 
	with $(X_0, Z_0)^\top \sim \mu^0$. The arguments of Section \ref{sec:ergodic} can be repeated to conclude that the invariant measure of $(X, Z)^\top$ admits a smooth density $\rho^0$ which satisfies \eqref{eq:FPsystem_homogenized}. Moreover, standard homogenization theory (see e.g. \cite[Chapter 3]{BLP78}) guarantees that $(X^\epl,Z^\epl)^\top \to (X,Z)^\top$ for $\epl \to 0$ in law as random variables with values in $\mathcal C^0([0, T]; \R^2)$. The Portmanteau theorem can be employed to conclude that the measure $\mu^\epl$ converges weakly to $\mu^0$ for $\epl \to 0$.
\end{proof}

@@

\corr{We can now prove the main result, namely the convergence of the filtering-based estimator to the drift coefficient of the homogenized equation.}

\begin{theorem}\label{thm:mainTheorem} Let the assumptions of Lemma \ref{lem:ergodicity} and Lemma \ref{lem:convMeasure} hold, and let $\widehat A^\epl_k(T)$ be defined in \eqref{eq:AHatMixed} \corr{with $\delta = \mathcal O(\epl^0)$}. If $\widetilde M$ is invertible, then
	\begin{equation}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A_k^\epl(T) = A, \quad \text{a.s.},
	\end{equation}
	where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
\end{theorem}

@@

\lim_{T \to \infty} I_1^\epl(T) = \widetilde {\mathcal M}_\epl^{-1} \E^{\rho^\epl} \left [ \frac{1}{\epl} p' \left ( \frac{X^\epl}{\epl} \right ) V'(Z^\epl) \right ], \quad \text{a.s.},
\widetilde {\mathcal M}_\epl \defeq \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ].

@@

	&= - \widetilde {\mathcal M}_\epl\alpha + \sigma \int_{\R} \int_{\R} V'(z) \phi^\epl(x) \psi^\epl(z) \partial_x R^\epl(x,z) \dd x \dd z.

@@

\lim_{T \to \infty} I_1^\epl(T) = -\alpha + \widetilde {\mathcal M}_\epl^{-1} \sigma \int_{\R} \int_{\R} V'(z) \phi^\epl(x) \psi^\epl(z) \partial_x R^\epl(x,z) \dd x \dd z, \quad \text{a.s.}

@@

	\lim_{T \to \infty} I_1^\epl(T) = -\alpha + \frac1\delta \widetilde {\mathcal M}_\epl^{-1} \E^{\rho^\epl}[(X^\epl - Z^\epl)^2 V''(Z^\epl)], \quad \text{a.s.}	

@@

\corr{Since $\delta = \mathcal O(\epl^0)$ we can now safely} pass to the limit as $\epl$ goes to zero and
\begin{equation} \label{eq:limit_estimator_2}
	\lim_{\epl \to 0} \lim_{T \to \infty} I_1^\epl(T) = -\alpha + \frac1\delta \widetilde {\mathcal M}_0^{-1} \E^{\rho^0}[(X - Z)^2 V''(Z)], \quad \text{a.s.},

@@

where the function $\rho^0$ is the density of the limit invariant distribution for $\epl \to 0$, as in Lemma \ref{lem:convMeasure}, and where
	\widetilde {\mathcal M}_0 \defeq \E^{\rho^0} [ V'(Z) \otimes V'(X) ].
Due to Corollary \ref{lem:FPMarginal_Hom}, we have

@@

	&=  \widetilde {\mathcal M}_0 A.

@@

	\lim_{T \to \infty} I_{2,1}^\epl(T) = \widetilde {\mathcal M}_\epl^{-1}\E^{\rho^\epl}\left[V'(Z^\epl) \otimes V'(Z^\epl)\right] \eqdef R^\epl,

@@

\corr{
\begin{remark}\label{rem:deltaPower} Let us remark that the condition $\delta = \mathcal O(\epl^0)$ is necessary to pass from \eqref{eq:for_zeta} to \eqref{eq:limit_estimator_2} but is not needed before \eqref{eq:for_zeta}. Moreover, the term $I_2^\epl(t)$ in the proof vanishes a.s. independently of $\epl$. Therefore, in the analysis of the case $\delta = \mathcal O(\epl^\zeta)$ it will be sufficient for unbiasedness to show that
		\lim_{\epl\to 0}\frac1\delta \widetilde {\mathcal M}_\epl^{-1} \E^{\rho^\epl}[(X^\epl - Z^\epl)^2 V''(Z^\epl)] = A,
	which is a non-trivial limit since $\delta \to 0$ for $\epl \to 0$.
\end{remark}
}

\subsection{Filtering at faster scales}\label{sec:Fast}

\corr{Remark for Assyr, Greg, Andrew: This section is all new, therefore it's ``red''. For easing the reading process, we write it here in black.}

We now consider the case of the filtering width $\delta = \mathcal O(\epl^\zeta)$, where $\zeta > 0$ will be specified in the following. In this case, the filtered process resembles more the original process $X^\epl$, as can be noted in Figure \ref{fig:filter}. Moreover, the techniques employed for proving Theorem \ref{thm:mainTheorem} can only be partly exploited, as highlighted by Remark \ref{rem:deltaPower}. In fact, in order to prove unbiasedness it is necessary to characterize precisely the difference between the processes $Z^\epl$ and $X^\epl$. A first characterization is given by the following Proposition, whose proof ought to be found in Appendix \ref{ap:ProofDistance}.

\begin{proposition} \label{prop:zeta} Let Assumption \ref{as:regularity} hold and $\epl,\delta>0$ be sufficiently small. Then, it holds for every $t > 0$ sufficiently big 
	\begin{equation}
	X_t^\epl - Z_t^\epl = \delta \beta^\epl_t + R(\epl,\delta),
	\end{equation}
	where the stochastic process $\beta_t^\epl$ is defined as
	\begin{equation}\label{eq:defBeta}
	\beta_t^\epl \defeq \sqrt{2\sigma} \int_0^t k(t-s)(1 + \Phi'(Y_s^\epl)) \dd W_s,
	\end{equation}
	where $\Phi$ is the solution of the cell problem \eqref{eq:CellProblem}, $W_s$ is the Brownian motion appearing in \eqref{eq:SDE_MS} and $Y_t^\epl = X_t^\epl / \epl$. Moreover, $\beta_t^\epl$ and the remainder $R(\epl,\delta)$ satisfy for every $p \geq 1$ the estimates
	\begin{equation} \label{eq:estimate_beta}
	\left( \E^{\phi^\epl} \abs{\beta_t^\epl}^p \right)^{1/p} \leq C \delta^{-1/2},
	\end{equation}
	and
	\begin{equation} \label{eq:remainder_zeta}
	\left( \E^{\phi^\epl} \left| R(\epl,\delta) \right|^p \right)^{1/p} \le C \left( \delta + \epl \right),
	\end{equation}
	where $C$ is independent of $\epl$ and $\delta$ and $\phi^\epl$ is the density of the invariant measure of $X^\epl$.
\end{proposition}

It is clear from the Proposition above that understanding the properties of the process $\beta_t^\epl$ is key to understanding the behavior of the difference between $X^\epl$ and $Z^\epl$. In particular, it is relevant to study the asymptotic properties of $\beta_t^\epl$ for long times

\begin{lemma} \label{lem:quartet}
	Let $\eta^\epl$ be the invariant measure of the quartet $(X^\epl, Y^\epl, Z^\epl, \beta^\epl)$. Then it holds
	\begin{equation}
	\delta \E^{\eta^\epl} \left[ (\beta^\epl)^2 V''(Z^\epl) \right] = \sigma \E^{\eta^\epl} [V''(Z^\epl) (1 + \Phi'(Y^\epl))^2] + \widetilde R(\epl,\delta),
	\end{equation}
	where the remainder $\widetilde R(\epl,\delta)$ satisfies
	\begin{equation} \label{eq:quartet_remainder}
	\left| \widetilde R(\epl,\delta) \right| \le C \left( \delta^{1/2} + \epl \right).
	\end{equation}
\end{lemma}

\begin{lemma} \label{lem:quartet_convergence}
	Let $\delta = \epl^\zeta$ with $\zeta > 0$, then it holds
	\begin{equation}
	\lim_{\epl \to 0} \sigma \E^{\eta^\epl} [V''(Z^\epl) (1 + \Phi'(Y^\epl))^2] = \Sigma \E^{\phi^0} [V''(X)].
	\end{equation}
\end{lemma}

\begin{theorem}\label{thm:mainTheorem_zeta} Let the assumptions of Lemma \ref{lem:ergodicity} and Lemma \ref{lem:convMeasure} hold, let $\widehat A^\epl_k(T)$ be defined in \eqref{eq:AHatMixed} and let $\delta = \epl^\zeta$ with $\zeta \in (0,2)$. If $\widetilde M$ is invertible, then
		\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A_k^\epl(T) = A, \quad \text{in law},
		\end{equation}
		where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
\end{theorem}
\begin{proof}
	Let us introduce the notation
	\begin{equation}
	\mathcal A^\epl(\delta) = \frac1\delta \widetilde{\mathcal M}_\epl^{-1} \E^{\rho^\epl}[(X^\epl - Z^\epl)^2 V''(Z^\epl)],
	\end{equation}
	where
	\begin{equation}
	\widetilde{\mathcal M}_\epl = \E^{\rho^\epl} [ V'(Z^\epl) \otimes V'(X^\epl) ].
	\end{equation}
	Then following the proof of Theorem \ref{thm:mainTheorem} we only need to show that if $\delta = \epl^\zeta$ with \corr{$\zeta \in (0,2)$} we have
	\begin{equation}
	\lim_{\epl \to 0} \mathcal A^\epl(\delta) = A, \quad \text{in law}.
	\end{equation}
	Using Lemma \ref{prop:zeta} and geometric ergodicity (Lemma \ref{lem:ergodicity}), we have the following equality
	\begin{equation}
	\begin{aligned}
	\mathcal A^\epl(\delta) &= \widetilde{\mathcal M}_\epl^{-1} \frac1\delta \lim_{t \to \infty} \E [(X_t^\epl - Z_t^\epl)^2 V''(Z_t^\epl)] \\
	&= \widetilde{\mathcal M}_\epl^{-1} \frac1\delta \lim_{t \to \infty} \E \left[ \left( \delta \beta_t^\epl + R(\epl,\delta) \right)^2 V''(Z_t^\epl) \right] \\
	&\eqdef \widetilde{\mathcal M}_\epl^{-1} \lim_{t \to \infty} \left( J_1^\epl(t) + J_2^\epl(t) + J_3^\epl(t) \right),
	\end{aligned}
	\end{equation}
	where and $\E$ denotes the expectation with respect to the Wiener measure and
	\begin{equation}
	\begin{aligned}
	J_1^\epl(t) &= \delta \E \left[ (\beta_t^\epl)^2 V''(Z_t^\epl) \right], \\
	J_2^\epl(t) &= 2 \E \left[ \beta_t^\epl R(\epl,\delta) V''(Z_t^\epl) \right], \\
	J_3^\epl(t) &= \frac1\delta \E\left[ R(\epl,\delta)^2 V''(Z_t^\epl) \right].
	\end{aligned}
	\end{equation}
	Let us consider the three terms separately. First, by geometric ergodicity and applying Lemma \ref{lem:quartet}, we get
	\begin{equation}
	\lim_{t \to \infty} J_1^\epl(t) = \delta \E^{\eta^\epl} \left[ (\beta^\epl)^2 V''(Z^\epl) \right] = \sigma \E^{\eta^\epl} [V''(Z^\epl) (1 + \Phi'(Y^\epl))^2] + \widetilde R(\epl,\delta),
	\end{equation}
	and, by Lemma \ref{lem:quartet_convergence} and due to \eqref{eq:quartet_remainder}, we obtain
	\begin{equation}
	\lim_{\epl \to 0} \lim_{t \to \infty} J_1^\epl(t) = \Sigma \E^{\rho^0} [V''(X)].
	\end{equation}
	Moreover, due to \eqref{eq:estimate_beta} and \eqref{eq:remainder_zeta}, we have the following bounds for $J_2^\epl(t)$ and $J_3^\epl(t)$
	\begin{equation}
	\begin{aligned}
	\abs{J_2^\epl(t)} &\le C \left( \delta^{1/2} + \epl \delta^{-1/2} \right), \\
	\abs{J_3^\epl(t)} &\le C \left( \delta + \epl^2 \delta^{-1} \right).
	\end{aligned}
	\end{equation}
	Furthermore, by Lemma \ref{lem:distanceMandTildeM} and weak convergence of the invariant measure, we have
	\begin{equation}
	\lim_{\epl \to 0} \widetilde{\mathcal M}_\epl^{-1} = \E^{\rho^0} [V'(X) \otimes V'(X) ] \eqdef \mathcal M_0.
	\end{equation}
	Therefore
	\begin{equation}
	\lim_{\epl \to 0} \mathcal A^\epl(\delta) = \Sigma \E^{\rho^0} [V'(X) \otimes V'(X) ]^{-1} \E^{\rho^0} [V''(X)],
	\end{equation}
	and, finally, an integration by parts yields
	\begin{equation}
	\lim_{\epl \to 0} \mathcal A^\epl(\delta) = \Sigma \E^{\rho^0} [V'(X) \otimes V'(X) ]^{-1} \frac1\Sigma \E^{\rho^0} [V'(X) \otimes V'(X) ] A = A,
	\end{equation}
	which implies the desired result.
\end{proof}

@@

\section{Proofs of Sections \ref{sec:ergodic}}\label{ap:Proofs}

@@

	which concludes the proof. \qed

\section{Proofs of Section \ref{sec:Fast}} \label{ap:zeta}

\subsection{Preliminary estimates}

Corollary \ref{cor:distanceZandX} is a direct consequence of Lemma \ref{prop:zeta} and provides a rough estimate of the difference between the trajectories $X_t^\epl$ and $Z_t^\epl$ when they are at stationarity.

\begin{corollary} \label{cor:distanceZandX} Under Assumption \ref{as:regularity}, let the couple $(X^\epl, Z^\epl)^\top$ be distributed as its invariant measure $\mu^\epl$. Then, if $\delta \leq 1$, it holds for any $p \geq 1$
	\begin{equation}
	\left(\E^{\rho^\epl} \abs{X^\epl - Z^\epl}^p\right)^{1/p} \leq C\left(\epl + \delta^{1/2}\right),
	\end{equation}
	for a constant $C > 0$ independent of $\epl$ and $\delta$.
\end{corollary}
\begin{proof}
	Let $p \geq 1$, then due to Lemma \ref{prop:zeta} there exists a constant $C > 0$ depending only on $p$ such that
	\begin{equation}
	\E^{\phi^\epl}\abs{X_t^\epl - Z_t^\epl}^p \leq C\left(\epl^p + \delta^{p/2}\right).
	\end{equation}
	Let us now remark that this result holds for $X_t^\epl$ being at stationarity and for $Z_t^\epl$ be its filtered process, and not for a couple $(X^\epl, Z^\epl)^\top \sim \mu^\epl$. In order to conclude, we remark that due to Lemma \ref{lem:ergodicity} we have for all $t \geq 0$ and for a constant $C$ depending on the initial condition and $p$  
	\begin{equation}
	\E^{\rho^\epl} \abs{X^\epl - Z^\epl}^p \leq \E^{\phi^\epl}\abs{X_t^\epl - Z_t^\epl}^p + Ce^{-\lambda t},
	\end{equation}
	which, for $t$ sufficiently big, yields the desired result.
\end{proof}
